---
title: "Reproducible Research Peer Grade Assessment Project 1"
author: "Sandip Mondal"
output:
  html_document:
    keep_md: yes
  pdf_document: default
  word_document: default
---  
  
Data, i.e. activity.zip is downloadef from [this GitHub repository](http://github.com/rdpeng/RepData_PeerAssessment1) on March 19, 2018 and kept in R working directory
  
  
  
## Loading and preprocessing the data
  
  
  
### Unzipping the data to get "activity.csv"
```{r unzip, echo=TRUE,results='hide', cache=TRUE}
unzip("activity.zip")
```
  
  
### Reading "activity.csv" and displaying first few sample rows
```{r readcsv, echo=TRUE, cache=TRUE}
Acts <- read.csv("activity.csv")
head(Acts)
```
  
  
  
## What is mean total number of steps taken per day?
  
  
  
### Calculating total number of steps taken per day
```{r calc_steps, echo=TRUE, cache=TRUE}
totSteps <- aggregate(steps ~ date, data = Acts, sum, na.rm = TRUE)
print(totSteps)
```
  
  
  
### Creating histogram of the total number of steps taken each day
```{r create_tot_steps_hist, echo=TRUE, cache=TRUE}
hist(totSteps$steps)
```
  
  
  
### Calculating mean of the total number of steps taken per day
```{r calc_mean, echo=TRUE, cache=TRUE}
mean(totSteps$steps)
```
  
  
  
### Calculating median of the total number of steps taken per day
```{r calc_median, echo=TRUE, cache=TRUE}
median(totSteps$steps)
```
  
  
  
## What is the average daily activity pattern?
  
  
  
### Makeing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r time_series_plot, echo=TRUE, cache=TRUE}
intv_stpAvg <- aggregate(steps ~ interval, data = Acts, mean, na.rm=TRUE)
plot(steps ~ interval, data = intv_stpAvg, type = "l")
```
  
  
  
### Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r max_avg, echo=TRUE, cache=TRUE}
intv_stpAvg[which.max(intv_stpAvg$steps), ]$interval
```
  
As per dataset being considered here 835 interval contains the maximum number of steps average across all the days.
  
  
  
## Imputing missing values
  
  
 
### Calculating and reporting the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r no_of_NA, echo=TRUE, cache=TRUE}
sum(is.na(Acts))
```
  
As per dataset being considered total number rows with missing value is 2304.
  
  
 
### Missing values are being replaced by mean values for that days
```{r NA_cleaning, echo=TRUE, results='hide', cache=TRUE}
acts_noNA <- Acts

acts_noNA$day_name <- weekdays(as.POSIXlt(Acts$date), abbreviate = FALSE)

for (i in 1:nrow(acts_noNA)) {
    if(is.na(acts_noNA[i, ]$steps)) {
        acts_noNA[i, ]$steps <-mean(acts_noNA[acts_noNA$day_name == acts_noNA[i, ]$day_name,]$steps, na.rm = TRUE)
    }
}
```
  
  
  
### Creating histogram of the total number of steps taken each day
```{r create_tot_steps_hist_noNA, echo=TRUE, cache=TRUE}
totSteps_noNA <- aggregate(steps ~ date, data = acts_noNA, sum)
hist(totSteps_noNA$steps)
```
  
  
  
### Calculating mean of the total number of steps taken per day
```{r calc_mean_noNA, echo=TRUE, cache=TRUE}
mean(totSteps_noNA$steps)
```
  
  
  
### Calculating median of the total number of steps taken per day
```{r calc_median_noNA, echo=TRUE, cache=TRUE}
median(totSteps_noNA$steps)
```
  
Imputing missing data on the estimates of the total daily number of steps is **making the result different**.
  
  
  
## Are there differences in activity patterns between weekdays and weekends?
  
  
### Creating a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day
```{r create_factor,  echo=TRUE, cache=TRUE}
acts_noNA$week_or_weekend <- ifelse(acts_noNA$day_name %in% c("Saturday", "Sunday"), "weekend", "weekday")
```
  
  
  
### Making a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r create_panel_plt,  echo=TRUE, cache=TRUE}
library(lattice)
acts_noNA <- aggregate(steps ~ interval + week_or_weekend, acts_noNA, mean)
xyplot(steps ~ interval | factor(week_or_weekend), data = acts_noNA, aspect = 1/2, 
       type = "l")
```

